A community-maintained standard library of population genetic models

### 2020-04-29 response-to-reviewers.md

> 1) Ensuring that the language about the models & parameters conveys the correct sense of uncertainty. Obviously the reviewers and the authors know that these just represent best guesses at the moment, but as this platform catches on we don't want these numbers & models to be seen as gospel.

We fully agree that the language in the original manuscript was not careful enough in conveying uncertainty about the models and parameters we discuss.
To address this, we have made edited the text throughout the manuscript trying to better convey this uncertainty.
See our response to Reviewer 1 below for examples.
We have also added a bit more to the discussion (below "next steps") on this point.

> 2) We would like to see the SLIM integration demonstrated with an application that msprime doesn't cover. This could be, for example, a figure of the average value some summary statistics surrounding a selective under the human demographic model. The reviewers didn't want this to be a lot of work, e.g., it would be outside of the scope of the current paper to demonstrate the power of a range of selection methods. However, we felt that a simple demonstration of the functionality would substantially increase the scenarios easily available and imaginable by readers of the paper.

We would like to see this also and agree that it would be of great interest.
However, we felt that performing adequate simulations with selection would require
a substantial amount of additional work and would significantly expand the scope of
the manuscript.
Furthermore, discussing selection would distract from the current relatively simple focus
on inference of demographic history.
We definitely intend to include selection (and expand the available generic models) in future work,
as we now discuss in greater detail in the Discussion ("next steps").

Consequently, we decided to add an example for possible use of SLiM in neutral simulations (see "simulation engines" section in the results and Supplementary Figures S1 and S2).
Our focus there was to show that under neutral simulations, SLiM produced data that were consistent with the coalescent simulations of msprime.
To expand on this fundamental point, we also examined the influence of population size down-scaling, which is a common practice used in forward simulations.
While this is by no means a comprehensive investigation into this issue, we do believe that it demonstrates the use of SLiM in future applications that will also simulate selection.

@@

> 3) We felt that the discussion would benefit from a discussion of the details not yet incorporated into the platform, e.g. sex-specific recombination, gene conversion, assembly error, structural variants etc. Such an addition might spur future development, and also give the authors an opportunity to discuss general ways forward on these issues.

Good idea -
we have expanded the section of the discussion ("next steps")
with more detail on our future plans.
We're trying to walk the line between being forward-looking and promising too much.

> 4) We would be interested to see more of a discussion of the strengths and weakness of the various demographic inference methods.

We have added an additional paragraph to the discussion exploring various factors that may affect choice of inference method.
These include the data required by the method, the type of model, and software engineering.
We also provide a qualitative comparison between the performance of methods, as shown in our limited analysis.
We believe that a more comprehensive comparison between methods is beyond the scope of the current manuscript, which focuses on the resource itself, rather than its application.

@@

The second passage now reads:
" Firstly, the library defines some basic information about our current understanding of each 
species' genome, including information about chromosome
lengths, average mutation rate estimates, and generation times.
We also provide access to detailed empirical information such as inferred genetic maps,
which model observed heterogeneity in recombination rate along chromosomes."

(The reviewer's suggestion is good, but we think that "current understanding"
conveys roughly the same meaning as "the latest reference genome" in less words,
and without making promises about genome assemblies.)

> b. On a related note, there is wording on page 4, regarding ensuring "implemented models are accurate" - I think what is meant is that the "implemented models are faithful to the source publications from which they derive". As the second half of this paper makes clear, because of errors in inference, many of these models will not be accurate, in the sense of representing the true history of the species.

We made the suggested change:
"Importantly, we developed rigorous quality control methods to ensure that we have
correctly implemented the models as described in their original publication
and provide documented methods for others to contribute new models."

@@

> 2. It would be wonderful to have a comments section for the models that would be either curated by a set of editors or crowd-sourced. I say this because over time, models will proliferate, and some will come to be regarded as out-of-date. One can imagine those approaching the field afresh will be overwhelmed by the possible selections and possible implement models that become outdated. If the goal is standardization, how do we communicate that standards change? A comment system (or even star-rating system?) may be wise to implement now. Assuming it can be layered on top of the existing structure, it may be enough for this publication to note this as a future challenge that needs development/addressing.

Thank you for this forward-thinking comment.
We considered various ways to do this
(e.g., by enabling the wiki associated with the github repository:
\url{https://github.com/popsim-consortium/stdpopsim/issues/415}),
but ended up deciding that (a) it's not yet clear what the right way to do this is,
and (b) it'll be significant work to do right that we aren't prepared to do at the moment.
Since we haven't encountered the problem of out of date models yet,
we feel happy about postponing this until the problem arises.

> 3. In terms of the maturity of the examples developed for the initial release, I would have liked to see at least one simulation model with a selective sweep, one with background selection, and one with spatial stepping-stone structure. Each of these would be helpful test cases to implement to be sure that the existing catalog framework has the breadth/flexibility necessary to accommodate future use cases. I do not think this is a requirement for publication, but it would add great value to this initial release of the resource.

As mentioned in our response to the editor's comments, we agree that this is an important avenue to pursue in the near future.
However, performing adequate simulations with selection would require
a substantial amount of additional work and would likely distract from the current relatively simple focus
on inference of demographic history.
We definitely intend to include selection (and expand the available generic models) in future work,
as we now discuss in greater detail in the Discussion ("next steps").

@@

> 4. Page 15. The approach of masking "low-recombination" portions of the chromosomes seems like an incomplete/indirect attempt to model the inherent limitations of sequencing to an "accessible" genome.
> a. Shouldn't the approach instead be to drop "low complexity" regions (e.g., as defined by an excessive number of "N"'s in the reference, low mapability scores, or via tools like RepeatMasker?). This part of the pipeline seems open to refinement.

Our initial motivation for masking was to reduce the overrepresentation of marginal trees with little to no recombination from biasing patterns of diversity in such a way that demographic inference methods would be misled. While we agree with the reviewer that different masking approaches might better reflect the masking done on real genomic data, at this time the best way to mask remains an open question, and we feel that a nuanced analysis of masking is outside the scope of this paper. For these reasons, we feel that masking using a simple recombination rate threshold is the most appropriate approach. We agree that there may indeed be better ways to mask, which is why we allow users to mask their simulations as they see fit.

@@

> b. Are the "masks" a separate configuration file for the simulations? It seems that it would be preferable for them to be separate from the recombination rate files - right now it reads as if the mask applied is a function of the genetic map file, but this seems too inflexible for users who prefer an alternative approach to masking.

Mask files are not currently a component of `stdpopsim` proper, rather they were implemented separately from running `stdpopsim`, for the sole purpose of comparing demographic inference methods. Users who download `stdpopsim` will always be simulating raw and unmasked tree sequence files, to which they can apply any variety of masks *post hoc*, if they so choose.

@@

> 6. Page 6. Typo in "time or writing". Change to "time of writing"

Thanks for spotting this typo! We corrected it.

> 7. Page 8. I like the approach of showing inverse coalescent rates, but it would be instructive to practitioners and useful for teaching, if in all figures you showed both the census sizes (e.g. with light grey dashed lines) and inverse rates (e.g. with dark grey solid lines as now in Fig 2).

First, we note that we decided to avoid using "census size", and simply refer to it as the "simulated population size", as we realized the use of "census" here could be misleading (see below for further discussion). Regarding the main point raised by the reviewer, we agree that in the single-population analyses, it would be instructive to show both the simulated (census) population size and the inverse coalescence rates. We thus plotted the simulated population sizes against the inverse coalescence rate for the model used in Figure 2, and found that they were near identical. As a result, we decided to show this comparison in Supplementary Figure S1 and discuss this issue in the text without adding the simulated population sizes to the main figures (as they didn't appear to add much). Note that in the multi-population analyses, the inverse coalescence rate is less appropriate, because some of the inference methods used in these analyses model gene flow, and should therefore be able to recover the true simulated population sizes. We now clarify this in the text as well.

@@

> 8. Figure 4.
> a. Confusing how in panels A and B time moves left to right, and in panel C time moves right to left.
> b. Also, the figure would be improved by showing with vertical lines the timing of the CEU/YRI divergence in the true model. (similar comments hold for Figures S4 and S5)

We agree that the orientation of this figure is confusing, and so have changed it as suggested. In addition, we had previously plotted divergence times here as vertical lines, however we felt that plotting it this way made the figure hard to read since there was a great deal of overlap between these lines and the N(t) lines on the figure, which was was especially difficult to read when including multiple replicates. Thus, we decided to depict the divergence times (simulated and inferred) in a separate panel in the bottom of the figure, to improve readability.

@@

Thank YOU! We hope you are right.

## Minor comments:

> The authors mention the implementation of the SLiM package into stdpopsim, but I was unable to find any details of this either in the manuscript or on the readthedocs page. If this was indeed implemented, I would also like to see the authors add some tutorials/results of analyses based on SLiM simulated data.

As mentioned in our response to the editor's comment, we added a demonstration of the use of SLiM to the results (see "Simulation engines" section and Supplementary Figures S1 and S2).
This new analysis uses SLiM in neutral simulations to validate that it produces data that is consistent with data simulated under the coalescent.
Furthermore, we provide a rudamentary examination of the effects of population size down-scaling, which is a common practice used in forward simulations such as SLiM.

@@

Thanks! We hope it is useful to you and your students.

## Comments

> 1) The authors mention that SLiM can be used as an alternative backend, which would presumably allow for simulations with selection. Although I don't think an extensive comparison of selection methods is necessary for this paper, it would be ideal if the authors can give some idea of how this would work (example command line, etc). There are also a myriad of methods for detecting/quantifying selection, and these simulations are not consistent either.

As mentioned in our response to the editor's comment, we added a demonstration of the use of SLiM to the results (see "Simulation engines" section and Supplementary Figures S1 and S2).
See also our response to related comments made by Reviewers #1 and #2.

@@

> 2) I like the inclusion of the "zigzag" history, as well as generic piecewise constant models and IM models (page 7). I wonder if these could be included in a separate section (not organism specific) in the documentation and software (and then in Table 1). Right now the zigzag model is under humans in the catalog.

This is an interesting idea. We did discuss adding a suite of generic models to stdpopsim that do not necessarily capture demographic parameters from a certain species,
but do reflect general trends observed in inference. The "zigzag" model could definitely be one of these.
This did not make it to this release, but we will note this for the next one.
the reason why the zigzag model is currently defined as a human model is that its
effective population size values are taken from (or at least inspired by) values inferred from human genomes.

@@

We made the suggested changes.

> -Page 12, first paragraph: include a reference where people could learn more about the effective population size vs. census size, random-mating models, etc

We thank the reviewer for this suggestion.
However, we have realized that our use of "census size" here to describe the "true" simulated population size was a bit misleading,
as our intention was not to characterize these as actual "census" (as opposed to "effective") population sizes.
To avoid this confusion, we have changed the manuscript to now describe them as the "simulated population size". For example, this section now reads:

"However, since the model we use for inference also
allows gene flow between populations, we directly compare estimated effective population sizes
to the values used in simulations (black line in
Figure 4C) and not the inverse coalescence rates."

@@

> -Page 12: S7, S4, S6 are mentioned in a different order than they appear in the supplement (I understand why but maybe revisit this part)

We reordered the figures to make sure they were referenced in the same order they appear in the supplement.

> -Page 13: "overfit the YRI history" do you mean overfit in a technical sense?

We meant to say that the inferred history has more population size changes
than the truth did;
we've change this to "overcomplicate the simple YRI history".

### 2020-04-29 popSimManu.tex
IG edits
@grahamgower edits

.

The \stdpopsim library has been developed by the PopSim Consortium using a
distributed open source model, with strong procedures in place
to continue its growth and maintain its quality.
Importantly, we have rigorous quality control methods to ensure that we have
correctly implemented the models as described in their original publication
and have documented methods for others to contribute new models.
We invite new collaborators to join our community:
those interested should visit our developer documentation at
\url{https://stdpopsim.readthedocs.io/en/latest/development.html}.
Below we describe the resource and give
examples of how it can be used to benchmark demographic inference methods.

@@

Coalescent simulations, while highly efficient, are limited in their ability to model
complex selection scenarios, such as recurrent sweeps and background selection. For these reasons
we have also implemented the forward-time simulator, \texttt{SLiM} \citep{haller2019tree,haller2019slim},
as an alternative backend engine to \texttt{stdpopsim}, allowing for the simulation of processes that cannot be modeled under
the coalescent. However, as forward-time simulators explicitly model all individuals in a population, simulating large
population sizes can be highly demanding of computational resources. One solution to this obstacle is to
rescale the population size, \textit{N}, such that population-dependent parameters (e.g. \textit{N$\mu$}, \textit{Nr},
and \textit{Ns}) remain unchanged \citep[see the \texttt{SLiM} manual and][]{uricchio2014robust}.

We validated our implementation of the \texttt{SLiM} engine by comparing
estimates of several population genetic summary statistics for neutral simulations generated by both \texttt{SLiM}
and \texttt{msprime} under different factors of population-size rescaling
(Figures \ref{fig:slim_val_map} and \ref{fig:slim_val_nomap}). We show that rescaling by up to a factor
of $50$ ($Q=50$ in \texttt{SLiM}) produces patterns of both diversity and linkage disequilibrium that are
qualitatively similar to those observed under the coalescent (i.e. \texttt{msprime}), when both are simulated under a recombination map
(Figure \ref{fig:slim_val_map}). It is important to note that rescaling \textit{N} is not, in actuality, the equivalent to simulating
all individuals in the population, and that differences between the two methods may be subtle and difficult to interpret.
To emphasize this point, we show deviations between patterns of diversity generated by \texttt{SLiM} and those generated by
\texttt{msprime} engines at $Q=50$, when simulating under a uniform recombination rate (Figure \ref{fig:slim_val_nomap}).
The causes of these differences remain unclear.
While an in-depth investigation of the effects of population-size rescaling under selection
scenarios is outside the scope of this report, we are enthusiastic about the potential for \texttt{stdpopsim} and the
\texttt{SLiM} engine to facilitate addressing such pertinent questions in the near future.

\subsection*{The Species Catalog}
The central feature of \stdpopsim is the species catalog, a systematic organization
of the key quantitative data needed to simulate a given species. These include a description
of the genome assembly, information about mutation rate estimates, inferred recombination rate(s), and generation time
in addition to a series of demographic models that have been estimated for that organism.

@@

@@

We start by comparing popular methods for estimating
population size histories of single populations and subsequently
show simple examples of multi-population inference.
To reproducibly evaluate and compare the performance of inference methods, we developed
workflows using \texttt{snakemake} \citep{koster2012snakemake},
available from \url{https://github.com/popsim-consortium/analysis},
that allow efficient computing in multicore or cluster environments.

For single-population population size histories, we compared
\MSMC~\citep{schiffels2014inferring}, \smcpp~\citep{terhorst2017robust}, and
\stairwayplot~\citep{liu2015exploring}
 on simulated genomes sampled from a single population,
in a number of the demographic models described above.
Our workflow generates $R$ replicates of $C$ chromosomes,
producing $n$ samples in each of a total of $R \times C$ simulations for each demographic model.
After simulation,
the workflow prepares input files for each inference method
by grouping all $n \times R \times C$ simulated chromosomes
into a single file.
Each file is then converted into an input file appropriate for each inference method
(so, all inference methods run on the same simulation replicates).
Each of the inference programs are then run in parallel, followed by plotting of
estimates of population size history from each program.

@@

Figure \ref{fig:n_t_ragsdale} presents the results
from simulations under OutOfAfricaArchaicAdmixture\_5R19,
a model of human migration out of Africa that includes archaic admixture
\citep{ragsdale2019models}, along with an empirical genetic map. In each column of this figure
we show the inferred population size history (denoted $N(t)$)
from samples taken from each of the three extant populations in the model.
In each row we show comparisons among the methods (including two sample sizes for \MSMC).
Blue lines show estimates from each of three replicate whole genome simulations.
When fitting single-population models to data simulated for multiple
populations with migration, however, we do not necessarily expect the inference methods to recover the
``true'' simulated population size due to model misspecification.
Instead, many methods work by matching coalescence time distributions, and a single-population model 
with varying population size can be fit to data generated from a more complex model including migration. 
In this case, the coalescence time distributions
will exactly match, and the coalescence rate is the inverse of the effective population size.
For this reason,
we plot the inverse coalescence rates (black lines) calculated analytically in \texttt{msprime}
as a benchmark of the ``truth'' in the single-population inference setting. However,
in this case inverse coalescence rates are very close to census population sizes
(Figure \ref{fig:census_vs_IRC}).
See the Appendix for a precise definition and description of the inverse coalescence rate computation.
While there is variation in accuracy among methods, populations, and individual replicates,
the methods generally infer back a good estimate
of the true effective population sizes of the simulations --
inferred values are mostly within a factor of two of the truth,
and many of the methods infer a bottleneck at the correct time.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.7\linewidth]{display_items/homo_sapiens_two_pop_comp.pdf}
\caption{\textbf{Parameters estimated using a multi-population human model}.
Here we show estimates of $N(t)$ inferred using \dadi, \fastsimcoal, and \smcpp. \textbf{(A)} Data were generated by simulating
replicate human genomes under the OutOfAfrica\_3G09 model and using the HapMapII\_GRCh37 genetic map
inferred in \cite{international2007second}. \textbf{(B)} For \dadi and \fastsimcoal we show parameters inferred
by fitting the depicted IM model, which includes population sizes, migration rates, and a split
time between CEU and YRI samples. \textbf{(C)} Population size estimates for each population (rows)
from \dadi, \fastsimcoal, and \smcpp (columns).
In shades of blue we show $N(t)$ trajectories estimated from each simulation,
and in black simulated population sizes for the respective population.
The population split date, $T_{DIV}$, is shown at
the bottom, with a common $x$-axis to the population size panels.}
\label{fig:IM_popn_human}
\end{center}
\end{figure}

\paragraph*{Multi-population demographic models.}
As \stdpopsim implements multi-population demographic models, we also
explored parameter estimation of population divergence parameters. In particular,
we simulated data under multi-population models for humans and \textit{D.~melanogaster}
and then inferred  parameters using \dadi, \fastsimcoal, and \smcpp.
For simplicity, we conducted inference in \dadi and \fastsimcoal by fitting an isolation with migration (IM) model
with constant population sizes and bi-directional migration \citep{hey2004im}.
Our motivation for fitting this simple
IM model was to mimic the typical approach of two population inference on empirical
data, where the user is not aware of the `true' underlying demography and the inference
model is often misspecified.
For human models with more than two populations \citep[e.g.,][]{gutenkunst2009inferring}
this means that we are inferring parameters for a model that does
not match the model from which the data were generated (Figures
\ref{fig:IM_popn_human}A and B). However, because model misspecification
here is less severe compared to the single-population case, we compare
our inferred population sizes only to the simulated population sizes (black line in
Figure \ref{fig:IM_popn_human}C) and not the inverse coalescence rates.

@@

Again, we can compare between species and look at the performance of these methods
for a two-population model of \textit{D.~melanogaster}. Figure
\ref{fig:two_popn_fly} shows inference results using data simulated under
the OutOfAfrica\_2L06 model. This model includes an ancestral population in
Africa from which a European population splits off following a bottleneck, with no
post-divergence gene flow between the African and European population (Figure \ref{fig:two_popn_fly}A).
Here again, we find that \dadi and \fastsimcoal infer more consistent histories,
but but do not detect the brief bottleneck in Europe, due to inference model
not allowing for population size changes after the population split. In addition, \dadi
and \fastsimcoal both do reasonably well at correctly inferring the absence of
migration (Figure \ref{fig:dromel_mig_rates}). In contrast, the inferred
demographic parameters from \smcpp are more noisy, though in some cases better
capture the short bottleneck in the European population.

@@

One of our goals is to engage research communities studying other taxa,
so as to expand the resource to many more species.
Although we have included demographic models and recombination maps,
there are many biological processes that we do not model.
Some of these additions that we are enthusiastic to add are:
selection (including distributions of fitness effects, maps of functional elements, and single sweeps),
gene conversion, mutation models (rate heterogeneity),
more realistic demography (overlapping generations, separate sexes, mortality/fecundity schedules),
geographic population structure,
and downstream aspects of data quality (genotyping and mapping error).
Moreover, an in-depth investigation into the effects of population-size rescaling under many of the
above scenarios is warented, given our preliminary findings using neutral simulations
(Figures \ref{fig:slim_val_map} and \ref{fig:slim_val_nomap}).
These plans would not be possible without recent important work on SLiM \citep{haller2019slim},
as well as current work on \texttt{msprime}.
Some other important processes are more challenging to model with current simulation software,
such as
structural variation,
changing recombination maps over time,
transposable elements,
and context-dependent mutation.

% note: this is addressing reviewer comments about model uncertainty
In particular, although the included demographic models represent the current state of the art,
we anticipate the set of available models to expand
as new methods and new modeling frameworks are developed.
For instance, the current models all describe a small set of discrete, randomly mating populations,
which are likely good approximations for deep-time population history,
but may be less useful for methods describing dynamics of contemporary populations.
\Stdpopsim's framework is sufficiently general that more realistic population models
will be easily incorporated, once they are published.

@@

For the single-population task we ran human (HomSap) simulations
using a variety of models (see Table 1): OutOfAfricaArchaicAdmixture\_5R19, OutOfAfrica\_3G09,
a constant-sized generic model, and a two-epoch generic model where the
population size instantaneously decreased from $N=10^4$ to $N=10^3$
five hundred generations before the present. Each simulation
used the HapmapII\_GRCh37 genetic map. For \textit{D.~melanogaster}
we estimated $N(t)$ from an African sample simulated under the DroMel,
African3Epoch\_1S16 model using the Comeron2012\_dm6 map.
Finally, we ran simulations of \textit{A.~thaliana} genomes using the
AraTha African2Epoch\_1H18 model under the Salome2012\_TAIR7 map.
For each model, three replicate whole genomes were
simulated and the population size estimated from those data. In all cases we
set the sample size of the focal population to $N=50$ chromosomes.

Following simulation, low-recombination portions of chromosomes were masked
from the analysis in a manner that reflects the ``accessible" subset of sites
used in empirical population genomic studies \citep[e.g.,][]{danecek20111000,langley2012genomic}.
Specifically we masked all regions of \SI{1}{cM} or greater in the lowest 5th percentile of the empirical
distribution of recombination, regions which are nearly uniformly absent for
empirical analysis.

>>>>

The \stdpopsim library has been developed by the PopSim Consortium using a
distributed open source model, with strong procedures in place
to continue its growth and maintain its quality.
Importantly, we developed rigorous quality control methods to ensure that we have
correctly implemented the models as described in their original publication
and provide documented methods for others to contribute new models.
We invite new collaborators to join our community:
those interested should visit our developer documentation at
\url{https://stdpopsim.readthedocs.io/en/latest/development.html}.
Below we describe the resource and give
examples of how it can be used to benchmark demographic inference methods.

@@

\subsection*{The species Catalog}
The central feature of \stdpopsim is the species catalog, a systematic organization
of the key quantitative data needed to simulate a given species. These include a description
of the genome assembly, information about mutation rate estimates, inferred recombination rate(s), and generation time
in addition to a series of demographic models that have been estimated for that organism.

@@

 \subsection*{Simulation engines}

Coalescent simulations, while highly efficient, are limited in their ability to model
complex selection scenarios, such as recurrent sweeps and background selection. For these reasons
we have also implemented the forward-time simulator, \texttt{SLiM} \citep{haller2019tree,haller2019slim},
as an alternative backend engine to \texttt{stdpopsim}, allowing for the simulation of processes that cannot be modeled under
the coalescent. However, as forward-time simulators explicitly model all individuals in a population, simulating large
population sizes can be highly demanding of computational resources. One common practice used to address this challenge is to
scale down the population size used in the simulations (\textit{N}),
and then scale down times specified in the demographic model by the same factor, and scale up the mutation and recombination
rates and selection coefficients such that the scaled parameters, \textit{N$\mu$}, \textit{Nr},
and \textit{Ns}, remain unchanged \citep[see the \texttt{SLiM} manual and][]{uricchio2014robust}.
Such down-scaled simulations are not completely equivalent to simulating
all individuals in the population, and may lead to subtle differences.
However, since many sequence-based measures of population diversity remain unchanged when rescaling in this fashion,
this practice is very common when using forward-time simulators, such as \texttt{SLiM}.

We validated our implementation of the \texttt{SLiM} engine by comparing
estimates of several population genetic summary statistics for neutral simulations generated by both \texttt{SLiM}
and \texttt{msprime}.
Examples of this validation for different down-scaling factors are shown in
Figures \ref{fig:slim_val_map} and \ref{fig:slim_val_nomap}, which indicate that scaling down the simulated
population size by up to a factor
of $50$ ($Q=50$ in \texttt{SLiM}) produces patterns of both diversity and linkage disequilibrium that are
qualitatively similar to those observed under the coalescent (i.e. \texttt{msprime}), when both are simulated under a recombination map
(Figure \ref{fig:slim_val_map}).
We do observe a slight departure in values of Tajima's D when scaling down by a factor of $50$,
likely due to the subtle differences implied by down-scaling.
Interestingly, this effect of down-scaling is more prominent and extends to other sequence-based statistcs
when data sets are simulated under a uniform recombination rate (Figure \ref{fig:slim_val_nomap}).
Importantly, both comparisons validate the equivalence of \texttt{SLiM} and \texttt{msprime} when no down-scaling is applied,
and they suggest that down-scaling population sizes can significantly reduce running time at a fairly modest cost to accuracy.
However, since the effects of rescaling are not fully understood, scaling factors should be used as conservatively as practicable.
The framework provided by \texttt{stdpopsim} should enable a deeper inspection of this issue, which we expect will contribute to
best practices used when applying forward-time simulators, such as \texttt{SLiM}.

%Ilan: added this short subsection since this paragraph didn't quite fit any of the other subsections
\subsection*{Documentation and reproducability}

@@

We start by comparing popular methods for estimating
population size histories of single populations and subsequently
show simple examples of multi-population inference.
To reproducibly evaluate and compare the performance of inference methods, we developed
workflows using \texttt{snakemake} \citep{koster2012snakemake},
available from \url{https://github.com/popsim-consortium/analysis},
that allow efficient computing in multicore or cluster environments.
Our workflow generates $R$ replicates of $C$ chromosomes,
producing $n$ samples in each of a total of $R \times C$ simulations for each demographic model.
After simulation,
the workflow prepares input files for each inference method
by grouping all $n \times R \times C$ simulated chromosomes
into a single file.
Each file is then converted into an input file appropriate for each inference method
(such that all inference methods run on the same simulation replicates).
Each of the inference programs are then run in parallel, followed by plotting of
estimates of population size history from each program.

@@

\subsubsection*{ Single-population demographic models.}
For single-population demographic models, we compared
\MSMC~\citep{schiffels2014inferring}, \smcpp~\citep{terhorst2017robust}, and
\stairwayplot~\citep{liu2015exploring}
 on simulated genomes sampled from a single population,
in a number of the demographic models described above.
When fitting single-population models to data simulated for multiple
populations with migration, we do not necessarily expect the inference methods to recover the
actual simulated population size due to model misspecification.
Indeed, the best one may expect in such a scenario is to infer a model which accurately reflects the
coalescence time distribution of the simulated model. While in the inferred single-population model,
this distribution is only affected by the ancestral effective population sizes, in the simulated multi-population
model it is also influenced by migration between the target population and populations not analyzed in inference.
Since ancestral population sizes correspond to the inverse of the coalescence rates, by computing the
inverse coalescence rates from the simulated model, we obtain the best estimates
one may hope to infer by analyzing a single population. We thus analytically compute inverse coalescence
rates in \texttt{msprime} for each simulated model, and use them as a benchmark of the ``true'' effective population sizes.
See the Appendix for a precise definition and description of the inverse coalescence rate computation.

Figure \ref{fig:n_t_ragsdale} presents the results
from simulations under OutOfAfricaArchaicAdmixture\_5R19,
a model of human migration out of Africa that includes archaic admixture
\citep{ragsdale2019models}, along with an empirical genetic map. In each column of this figure
we show the inferred population size history (denoted $N(t)$)
from samples taken from each of the three extant populations in the model.
In each row we show comparisons among the methods (including two sample sizes for \MSMC).
Blue lines show estimates from each of three replicate whole genome simulations,
and black lines indicate the ``true" values depicted by the inverse coalescence rates.
The latter are included as an appropriate reference because
when fitting single-population models to data simulated for multiple
populations with migration, we do not necessarily expect the inference methods to recover the
actual sizes of the populations used in simulation due to model misspecification.
Instead, many methods work by matching coalescence time distributions, and a single-population model
with varying population size can be fit to data generated from a more complex model including migration.
In this case, the coalescence time distributions
will exactly match, and the coalescence rate is the inverse of the effective population size.
However, in this specific model,
the inverse coalescence rates are very close to the simulated population sizes
(Figure \ref{fig:census_vs_IRC}).
While there is variation in accuracy among methods, populations, and individual replicates,
the methods generally infer back a good estimate
of the true effective population sizes of the simulations --
inferred values are mostly within a factor of two of the truth,
and many of the methods infer a bottleneck at the correct time.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.7\linewidth]{display_items/homo_sapiens_two_pop_comp.pdf}
\caption{\textbf{Parameters estimated using a multi-population human model}.
Here we show estimates of $N(t)$ inferred using \dadi, \fastsimcoal, and \smcpp. \textbf{(A)} Data were generated by simulating
replicate human genomes under the OutOfAfrica\_3G09 model and using the HapMapII\_GRCh37 genetic map
inferred in \cite{international2007second}. \textbf{(B)} For \dadi and \fastsimcoal we show parameters inferred
by fitting the depicted IM model, which includes population sizes, migration rates, and a split
time between CEU and YRI samples. \textbf{(C)} Population size estimates for each population (rows)
from \dadi, \fastsimcoal, and \smcpp (columns).
In shades of blue we show $N(t)$ trajectories estimated from each simulation,
and in black simulated population sizes for the respective population.
The population split time, $T_{DIV}$, is shown at
the bottom (simulated value in black and inferred values in blue),
with a common $x$-axis to the population size panels.}
\label{fig:IM_popn_human}
\end{center}
\end{figure}

\subsubsection*{Multi-population demographic models.}
As \stdpopsim implements multi-population demographic models, we also
explored parameter estimation of population divergence parameters. In particular,
we simulated data under multi-population models for humans and \textit{D.~melanogaster}
and then inferred  parameters using \dadi, \fastsimcoal, and \smcpp.
For simplicity, we conducted inference in \dadi and \fastsimcoal by fitting an isolation with migration (IM) model
with constant population sizes and bi-directional migration \citep{hey2004im}.
Our motivation for fitting this simple
IM model was to mimic the typical approach of two population inference on empirical
data, where the user is not aware of the `true' underlying demography and the inference
model is often misspecified.
For human models with more than two populations \citep[e.g.,][]{gutenkunst2009inferring}
this means that we are inferring parameters for a model that does
not match the model from which the data were generated (Figures
\ref{fig:IM_popn_human}A and B). However, since the model we use for inference also
allows gene flow between populations, we directly compare estimated effective population sizes
to the values used in simulations (black line in Figure \ref{fig:IM_popn_human}C)
and not the inverse coalescence rates.

@@

Again, we can compare between species and look at the performance of these methods
for a two-population model of \textit{D.~melanogaster}. Figure
\ref{fig:two_popn_fly} shows inference results using data simulated under
the OutOfAfrica\_2L06 model. This model includes an ancestral population in
Africa from which a European population splits off following a bottleneck, with no
post-divergence gene flow between the African and European population (Figure \ref{fig:two_popn_fly}A).
Here again, we find that \dadi and \fastsimcoal infer more consistent histories,
but they do not detect the brief bottleneck in Europe, due to inference model
not allowing for population size changes after the population split. In addition, \dadi
and \fastsimcoal both do reasonably well at correctly inferring the absence of
migration (Figure \ref{fig:dromel_mig_rates}). In contrast, the inferred
demographic parameters from \smcpp are more noisy, though in some cases better
capture the short bottleneck in the European population.

@@

One of our goals is to engage research communities studying other taxa,
so as to expand the resource to many more species.
Although we have included demographic models and recombination maps,
there are many biological processes that we do not model.
Some of these additions that we are enthusiastic to add are:
selection (including distributions of fitness effects, maps of functional elements, and single sweeps),
gene conversion, mutation models (rate heterogeneity),
more realistic demography (overlapping generations, separate sexes, mortality/fecundity schedules),
geographic population structure,
and downstream aspects of data quality (genotyping and mapping error).
Moreover, an in-depth investigation into the effects of population-size rescaling under many of the
above scenarios is warranted, given our preliminary findings using neutral simulations
(Figures \ref{fig:slim_val_map} and \ref{fig:slim_val_nomap}).
These plans would not be possible without recent important work on SLiM \citep{haller2019slim},
as well as current work on \texttt{msprime}.
Some other important processes are more challenging to model with current simulation software,
such as
structural variation,
changing recombination maps over time,
transposable elements,
and context-dependent mutation.

% note: this is addressing reviewer comments about model uncertainty
In particular, although the included demographic models represent the current state of the art,
we anticipate the set of available models to expand
as new methods and new modeling frameworks are developed.
For instance, the current models all describe a small set of discrete, randomly mating populations,
which are likely good approximations for deep-time population history,
but may be less useful for methods describing dynamics of contemporary populations.
\Stdpopsim's framework is sufficiently general that more realistic population models
will be easily incorporated, once they are published.
Although the models in the catalog are state of the art, it is
important to acknowledge that these are current best estimates of parameter values
which will change over time. Additional complications such as changing
genome builds as assemblies improve must also incorporated.
As the selection of models and species grows, we hope to develop an improved
documentation system that incorporates crowd-sourced community input
to highlight the most appropriate models and genome details.

@@

For the single-population task we ran human (HomSap) simulations
using a variety of models (see Table 1): OutOfAfricaArchaicAdmixture\_5R19, OutOfAfrica\_3G09,
and a constant-sized generic model.
% ILAN REMOVED THIS APR 29, 2020 - NOT USED ANYMORE
%, and a two-epoch generic model where the population size instantaneously decreased from $N=10^4$ to $N=10^3$
% five hundred generations before the present.
Each simulation
used the HapmapII\_GRCh37 genetic map. For \textit{D.~melanogaster}
we estimated $N(t)$ from an African sample simulated under the DroMel,
African3Epoch\_1S16 model using the Comeron2012\_dm6 map.
Finally, we ran simulations of \textit{A.~thaliana} genomes using the
AraTha African2Epoch\_1H18 model under the Salome2012\_TAIR7 map.
For each model, three replicate whole genomes were
simulated and the population size estimated from those data. In all cases we
set the sample size of the focal population to $N=50$ chromosomes.

Following simulation, low-recombination portions of chromosomes were masked
from the analysis in a manner that reflects the ``accessible" subset of sites
used in empirical population genomic studies \citep[e.g.,][]{danecek20111000,langley2012genomic}.
Specifically we masked all regions of \SI{1}{cM} or greater in the lowest 5th percentile of the empirical
distribution of recombination, regions which are nearly uniformly absent for empirical analysis.
This approach to masking was chosen to prevent marginal trees with low or no recombination 
from biasing the comparisons of demographic inference methods. It should be noted that masking is not implemented 
within \texttt{stdpopsim} proper; tree sequences generated by \texttt{stdpopsim} are always raw and unmasked. 
This allows users the flexibility to implement masking approaches that are specific to their needs 
for downstream analysis.
